### 2023-05-21 Outline.tex
Initial Commit

\documentclass{article}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage{enumitem}

\begin{document}
    \begin{center} 
        \textbf{\LARGE{Linear and Nonlinear Filters}}\\
        \large\makebox[2.8in]{Alexander Skretting\hfill} 445 457\\
        \makebox[2.8in]{Jose Rigel Soeryo Soebandoro\hfill} 444 345
    \end{center}
    \large\begin{enumerate}[label=\Roman*.]
        \item \textbf{Introduction} \emph{Both}
            \begin{enumerate}[label=\arabic*.]
                \item Motivation: What problems prompts the use of filters?
                    \textit{(Unwanted noise, enhancement of image properties to extract information)}
                \item General difference between Linear and Nonlinear Filters
            \end{enumerate}
        \item \textbf{Linear Filters} \emph{Jose}
            \begin{enumerate}[label=\arabic*.]
                \item Formal definition of Linear Filter \textit{Mathematical formula/definition of a linear filter}
                \item Types of Linear Filters: \textit{Box, Gaussian, Derivative} General Formulas, implementation in code
                \item Use Cases \textit{ Noise Removal, Edge Detection, Image Sharpening}
                \item Convolution and Kernel \textit{Matrix Operation, neighboring pixels, normalization, separable convolution}
                    add implementation and code to show how matrices are used to pass pixels and transform them
                \item Optimization of sigma values \textit{How to determine the best sigma values,
                    what factors come into play to determine those values, what makes the best result and metrics of 'best' result}
            \end{enumerate}
        \item \textbf{Nonlinear Filters} \emph{Alexander}
            \begin{enumerate}[label=\arabic*.]
                \item Formal Defintion of Nonlinear Filter
                \item Types of Nonlinear Filters: \textit{Median Filtering, MinFilter, MaxFilter, MeanShift Filter, CellularAutomaton}
                \item Optimal Nonlinear Filtering \textit{Kushner–Stratonovich filtering}
                \item Typical use Cases \textit{Noise Removal}
            \end{enumerate}
        \item \textbf{Testing and Analysis} \emph{Both}
            \begin{enumerate}[label=\arabic*.]
                \item Methodology \textit{(Image sets to be used, metrics of analysis to determine quality)}
                \item Test scenarios \textit{Noise Removal, Image Sharpening, Edge Detection, Median Filtering, Morphological Filtering}
                \item Optimization of parameters based on each types of filter and how the optimization is determined.
                \item Results based on test cases \textit{(Graph, Data Table, etc)}
                \item Tests limitations \textit{What is and is not possible to be determined based on testing}
                \item Results discussion
            \end{enumerate}
        \item \textbf{Conclusion} \emph{Both}
            \begin{enumerate}[label=\arabic*.]
                \item Pros and Cons of each Types of Filters \textit{Is one filter objectively better? Do they work better for different scenarios?}
                \item Limitations of Filtering
                \item Further Analysis to be done \textit{What can supplement Filters for further use cases in image processing}
            \end{enumerate}
    \end{enumerate}

\end{document}