### 2023-06-19 Draft.tex
Started separable filtering

\RequirePackage[2020-02-02]{latexrelease}

% Template for seminar reports
% Computer Vision Group, Visual Computing Institute, RWTH Aachen University
\documentclass[twoside,a4paper,article]{combine}
\usepackage[utf8]{inputenc}
\usepackage{a4}
\usepackage{fancyhdr}   
%\usepackage{german}    % Uncomment this iff you're writing the report in German
\usepackage{makeidx}
\usepackage{color}
\usepackage{t1enc}		% German letters in the "\hyphenation" - command
\usepackage{latexsym}	% math symbols
\usepackage{amssymb}    % AMS symbol fonts for LaTeX.
\usepackage{graphicx}
\usepackage{pslatex}
\usepackage{ifthen}
\usepackage{booktabs}
\usepackage[T1]{fontenc}
\usepackage{pslatex}
\usepackage{psfrag}
\usepackage{subfigure}
\usepackage{url}
\usepackage{datetime}
\usepackage{xspace}
\usepackage{newtxmath}

\usepackage{float}
\usepackage{subcaption}

\newdateformat{monthyeardate}{\monthname[\THEMONTH] \THEYEAR}

% Do not change these sizes and do not add superfluous 
% pagebreaks to increase the page count.
\setlength{\oddsidemargin}{3.6pt}
\setlength{\evensidemargin}{22.6pt}
\setlength{\textwidth}{426.8pt}
\setlength{\textheight}{654.4pt}
\setlength{\headsep}{18pt}
\setlength{\headheight}{15pt}
\setlength{\topmargin}{-41.7pt}
\setlength{\topskip}{10pt}
\setlength{\footskip}{42pt}
\setlength{\parindent}{0pt}

\makeatletter
\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
\def\cf{\emph{c.f}\onedot} \def\Cf{\emph{C.f}\onedot}
\def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}
\def\wrt{w.r.t\onedot} \def\dof{d.o.f\onedot}
\def\etal{\emph{et al}\onedot}
\makeatother

% =========================================================================
\graphicspath{{pictures/}}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}

% =========================================================================
\begin{document}
\include{titlepage}

\begin{abstract}
% +++++++++++++++++++++++++
% Insert your Abstract here (one paragraph summary)
% +++++++++++++++++++++++++
\end{abstract}

\tableofcontents
\newpage
% =========================================================================

\section{Introduction}
\section{Linear Filters}
With \textbf{Linear Filtering}, or specifically linear spatial filtering, the function which is used to pass the image through
must be linear and shift invariant. The function \pmb{L} is considered linear if there exists two constants $a$ and $b$ for any
two input pixels $f_1(m, n)$ and $f_2(m, n)$ such that

\begin{align*}
    \implies & g_1(m, n) = \pmb{L} [f_1(m, n)] \land g_2(m, n) = \pmb{L} [f_2(m, n)] \\
    \implies & a \cdot g_1(m, n) + b \cdot g_2(m, n) = \pmb{L} [a \cdot f_1(m, n) + b \cdot f_2(m, n)]
\end{align*}

also called the \emph{superposition property of linear systems}\cite{BovikActon09}. Whereas a function is considered \emph{shift invariant} when
\begin{align*}
    \implies & g(m, n) = \pmb{L} [f(m, n)]\\
    \implies & g(m - p, n - q) =  \pmb{L} [f(m - p, n - q)]
\end{align*} $p$ and $q$ implies a spatial shift to both the input and output pixels\cite{BovikActon09}. An intuitive way to think
of these two properties is that the function has to behave the same throughout the entire picture.\smallskip

\subsection{Correlation and Convolution Filtering}
A common formula for linear filtering is the \emph{Correlation Filtering}\cite{BovikActon09}. 
\[
    g(i,j) = \sum_{l \in \mathscr{M}}\sum_{k \in \mathscr{N}}{f(i+k, j+l) \cdot h(k, l)}
\]
or commonly notated as $g = f \otimes h$.\smallskip

The desired output pixel $g(i, j)$, where $i$ and $j$ specify the
coordinates of it, is based on a $M \times N$ sized neighborhood, 
meaning not only does one pixel define an output pixel, but also a specified number
of its neighbors. The influence of each pixel in the neighborhood is defined by the filter coefficient $h(k, l)$,
also called its \emph{kernel} or \emph{mask}. \cite{Szeliski_2022}\smallskip

\[
\underbrace{
    \begin{bmatrix}
        128 & 34 & 123\\
        68 & 54 & 73 \\
        100 & 95 & 17
    \end{bmatrix}}_{\text{input neighborhood}}
\otimes
\underbrace{
    \begin{bmatrix}
        0.1 & 0.1 & 0.1\\
        0.1 & 0.2 & 0.1\\
        0.1 & 0.1 & 0.1
    \end{bmatrix}}_{\text{kernel}}
= 75
\]

As the above example with a $3 \times 3$ kernel, a total of 9 pixels is needed to calculate a single output pixel.\smallskip

Another common variant on the formula is having the signs of the offsets reversed. 
\[
    g = f \ast h
\]
\[
    g(i,j) = \sum_{l \in \mathscr{M}}\sum_{k \in \mathscr{N}}{f(i-k, j-l) \cdot h(k, l)}
\]
With this formula, $\ast$ is called the \emph{convolutional} operator, and the kernel $h$ is called the \emph{impulse response function}. 
An interesting note is that, when the kernel $h$ is convolved with an impulse signal $\delta$ (an image with 0 everywhere except the origin),
it reproduces the kernel itself $h \ast \delta = h$, whereas with correlational filtering, 
it produces the reflected signal (inverted signal in both dimensions).\cite{Szeliski_2022} In cases
where the kernel is symmetrical on both axis (e.g. box blur), the result of convolutional and correlational is the same.

\begin{minipage}{\textwidth}\begin{figure}[H]
    \centering
    \subfigure[correlation]{\includegraphics[width=0.19\textwidth]{img/correlation.jpg}}
    \subfigure[convolution]{\includegraphics[width=0.19\textwidth]{img/convolved}}
    \caption{Difference between Correlation and Convolution Filtering by translating 32 pixels}
\end{figure}\end{minipage}\smallskip

An apparent problem from neighborhood filtering is that on the edges, the neighbors simply does not exist in one or two directions 
(e.g. a $1000\times1000$ image passed through a $3\times3$ kernel would produce a $998\times998$ image). 
There are a couple method to alleviate the calculation of the nonexistent neighbors.

\subsection{Padding (Border Effects)}

\begin{minipage}{\textwidth}\begin{figure}[H]
    \centering
    \subfigure[zero]{\includegraphics[width=0.19\textwidth]{img/border_constant}}
    \subfigure[wrap]{\includegraphics[width=0.19\textwidth]{img/border_wrap}}
    \subfigure[clamp]{\includegraphics[width=0.19\textwidth]{img/border_clamp}}
    \subfigure[mirror]{\includegraphics[width=0.19\textwidth]{img/border_reflect}}
    \subfigure[extend]{\includegraphics[width=0.19\textwidth]{img/border_extend.jpg}}
    \caption{Different types of border padding}
    \label{fig:padding_borders}
\end{figure}\end{minipage}
\begin{itemize}
    \item \textbf{Zero:} Set all pixels outside the region to zero.
    \item \textbf{Constant:} This is similar to Zero, but instead a specific color is chosen to replace the out of bound pixels.
    \item \textbf{Clamp:} The edge pixels are repeated indefinitely.
    \item \textbf{(Cyclic) Wrap (Repeat or Tile):} The entire image is repeated similar to tiling it infinitely.
    \item \textbf{Mirror:} The image is mirrored accordingly on each of the edges.
    \item \textbf{Extend:} This is a combination of clamping and mirroring, where clamped image is subtracted with the mirrored image.
\end{itemize}

This of course not only affects the unseen pixels, but the effects can bleed into the pixels inside the frame.
\begin{minipage}{\textwidth}\begin{figure}[H]
    \centering
    \subfigure[zero]{\includegraphics[width=0.19\textwidth]{img/boxBlur_constant}}
    \subfigure[wrap]{\includegraphics[width=0.19\textwidth]{img/boxBlur_wrap}}
    \subfigure[clamp]{\includegraphics[width=0.19\textwidth]{img/boxBlur_clamp}}
    \subfigure[mirror]{\includegraphics[width=0.19\textwidth]{img/boxBlur_reflect}}
    \subfigure[extend]{\includegraphics[width=0.19\textwidth]{img/boxBlur_extend.jpg}}
    \caption{Effects of different paddings with box blur}
    \label{fig:boxBlur_borders}
\end{figure}\end{minipage}\smallskip

There are other ways to handle the borders such as two-phase duplication,
which is similar to clamp but alternates between the last 2 pixels.
A change in the filter function per individual basis can also be implemented, but it might not be practical,
as large amounts of additional logic may be needed\cite{BaileyAmbikumar18}.

\subsection{Separable Filtering}
The time complexity to calculate a single pixel can be said to be $O(M \cdot N)$ as it is dependent
on the size of the kernel. For each of the pixel in the kernel, it must be both multiplied and summed together
with its neighbors. This can be in practice sped up by separating the horizontal and vertical operations, hence performing
a one dimensional convolution horizontally, then one additional operation for the vertical convolution. This kernel in this case is the
outer product of two one dimensional kernels\cite{Szeliski_2022}. 
\[
    \pmb{K} = v \otimes h = 
    \begin{bmatrix}
        v_1h_1 & v_1h_2 & v_1 h_3 & \hdots & v_1h_n \\
        v_2h_1 & v_2h_2 & v_2 h_3 & \hdots & v_2h_n \\
        \vdots & \vdots & \vdots & \ddots & \vdots \\
        v_m h_1 & v_m h_2 & v_m h_3 & \hdots & v_m h_n \\
    \end{bmatrix}
\]

% =========================================================================
\bibliographystyle{alpha}
\bibliography{abbrev, Draft}
\end{document}