### 2023-06-25 main.tex

\documentclass{article}
\usepackage{biblatex}
\usepackage{graphicx}
\usepackage{float}
\addbibresource{prosem.bib}
\begin{document}
\printbibliography

\end{document}